---
title: "WednesdayNo.1"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    storyboard: true
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(tidymodels)
library(ggdendro)
library(heatmaply)
library(plotly)
library(shiny)
library(jsonlite) 
library(readr)
library(ggplot2)
library(tidyverse)
source("compmus.R")
compmus2025 <- read_csv("compmus2025.csv")
#      C     C#    D     Eb    E     F     F#    G     Ab    A     Bb    B
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )
```

## 1st attempt

## 2nd Attempt: Track-Level Analysis

# Feature Comparison {.tabset}

### Chromagram

## Track 1 {.column}
```{r}
"features/wednesday-w-1.json" |> 
  compmus_chroma(norm = "identity") |> 
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c("C", "C#|Db", "D", "D#|Eb", "E", "F", "F#|Gb", "G", "G#|Ab", "A", "A#|Bb", "B")
  ) +
  scale_fill_viridis_c(option = "G") + 
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()
```

> This track exhibits clear harmonic repetition, with strong vertical striping indicating repeated pitch patterns.  
> This kind of structure is closely related to groove and may contribute to its danceability.

## Track 2 {.column}
```{r}
"features/wednesday-w-2.json" |> 
  compmus_chroma(norm = "identity") |> 
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c("C", "C#|Db", "D", "D#|Eb", "E", "F", "F#|Gb", "G", "G#|Ab", "A", "A#|Bb", "B")
  ) +
  scale_fill_viridis_c(option = "G") + 
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()
```

> Compared to Track 1, this track shows more harmonic variation but less repetition.  
> It might be more expressive or fluid, but less predictable, which may slightly reduce its danceability.

---

### Cepstrogram

## Track 1 {.column}
```{r}
"features/wednesday-w-1.json" |> 
  compmus_mfccs(norm = "identity") |> 
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL
  ) +
  scale_fill_viridis_c(guide = "none") + 
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic()
```

## Track 2 {.column}
```{r}
"features/wednesday-w-2.json" |> 
  compmus_mfccs(norm = "identity") |> 
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL
  ) +
  scale_fill_viridis_c(guide = "none") + 
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic()
```

---

### Chroma-based SSM

## Track 1 {.column}
```{r}
"features/wednesday-w-1.json" |> 
  compmus_chroma(norm = "identity") |> 
  compmus_self_similarity(feature = pc, distance = "euclidean") |> 
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") + 
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()
```

## Track 2 {.column}
```{r}
"features/wednesday-w-2.json" |> 
  compmus_chroma(norm = "identity") |> 
  compmus_self_similarity(feature = pc, distance = "euclidean") |> 
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") + 
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()
```

---

### Timbre-based SSM

## Track 1 {.column}
```{r}
"features/wednesday-w-1.json" |> 
  compmus_mfccs(norm = "identity") |> 
  compmus_self_similarity(feature = mfcc, distance = "euclidean") |> 
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") + 
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()
```

## Track 2 {.column}
```{r}
"features/wednesday-w-2.json" |> 
  compmus_mfccs(norm = "identity") |> 
  compmus_self_similarity(feature = mfcc, distance = "euclidean") |> 
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") + 
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()
```

---

### Chordograms

## Track 2 {.column}
```{r}
"features/wednesday-w-2.json" |> 
  compmus_chroma(norm = "identity") |> 
  compmus_match_pitch_templates(
    key_templates,
    norm = "identity",
    distance = "cosine"
  ) |> 
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") + 
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()
```

## Track 1 {.column}
_(Chordogram not generated for Track 1 in this layout. Can be added if needed.)_

## 别的

```{r}
get_conf_mat <- function(fit) {
  outcome <- .get_tune_outcome_names(fit)
  fit |> 
    collect_predictions() |> 
    conf_mat(truth = outcome, estimate = .pred_class)
}  

get_pr <- function(fit) {
  fit |> 
    conf_mat_resampled() |> 
    group_by(Prediction) |> mutate(precision = Freq / sum(Freq)) |> 
    group_by(Truth) |> mutate(recall = Freq / sum(Freq)) |> 
    ungroup() |> filter(Prediction == Truth) |> 
    select(class = Prediction, precision, recall)
}  
```

```{r}
cluster_juice <-
  recipe(
    filename ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = compmus2025
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |> 
  # step_range(all_predictors()) |> 
  prep(compmus2025) |>
  juice() |>
  column_to_rownames("filename")
```

### Hierachical clustering

```{r}
compmus_dist <- dist(cluster_juice, method = "euclidean")
compmus_dist |> 
  hclust(method = "average") |> # Try single, average, and complete.
  dendro_data() |>
  ggdendrogram()
```

### Heatmaps

```{r}
heatmaply(
  cluster_juice,
  hclustfun = hclust,
  hclust_method = "average",  # Change for single, average, or complete linkage.
  dist_method = "euclidean"
)
```

### Classification

```{r}
compmus2025_filtered <- 
  compmus2025 |> filter(!is.na(ai)) |> 
  mutate(ai = factor(if_else(ai, "AI", "Non-AI")))
classification_recipe <-
  recipe(
    ai ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = compmus2025_filtered
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors())      # Converts to z-scores.
  # step_range(all_predictors())    # Sets range to [0, 1].
compmus_cv <- compmus2025_filtered |> vfold_cv(5)
#random forest
forest_model <-
  rand_forest() |>
  set_mode("classification") |> 
  set_engine("ranger", importance = "impurity")
indie_forest <- 
  workflow() |> 
  add_recipe(classification_recipe) |> 
  add_model(forest_model) |> 
  fit_resamples(
    compmus_cv, 
    control = control_resamples(save_pred = TRUE)
  )
indie_forest |> get_pr()
workflow() |> 
  add_recipe(classification_recipe) |> 
  add_model(forest_model) |> 
  fit(compmus2025_filtered) |> 
  pluck("fit", "fit", "fit") |>
  ranger::importance() |> 
  enframe() |> 
  mutate(name = fct_reorder(name, value)) |> 
  ggplot(aes(name, value)) + 
  geom_col() + 
  coord_flip() +
  theme_minimal() +
  labs(x = NULL, y = "Importance")
```

### filtered

```{r}
compmus2025_filtered |>
  ggplot(aes(x = valence, y = arousal, colour = ai, size = tempo)) +
  geom_point(alpha = 0.8) +
  scale_color_viridis_d() +
  labs(
    x = "Valence",
    y = "Arousal",
    size = "Tempo",
    colour = "AI"
  )
```

### How Mainstream Are we?

```{r}
boxplot(
  compmus2025$danceability,
  horizontal = TRUE, 
  col = "lightblue",
  main = "Approachability Boxplot",
  xlab = "Approachability Score",
  outpch = 16,
  outcol = "lightpink"
)
```

### Novelty

```{r}
"features/ahram-j-1.json" |>
  compmus_energy_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  scale_color_gradientn(colours =rainbow(10))+
  labs(x = "Time (s)", y = "Energy Novelty")
```

# assignment page

Column{.tabset}
--------

### Novelty

```{r}
"features/ahram-j-1.json" |>
  compmus_energy_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  scale_color_gradientn(colours =rainbow(10))+
  labs(x = "Time (s)", y = "Energy Novelty")
```

### Chromagram of Track 1 

```{r}
"features/wednesday-w-1.json" |>                           # Change the track
  compmus_chroma(norm = "identity") |>                 
  # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(option = "G") +               
  # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()  
```

### Cepstrogram 

```{r}
"features/wednesday-w-1.json" |>                           #Change the track
  compmus_mfccs(norm = "identity") |>                  #Change the norm
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +               #Change the colours?
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic()                                      
```

### Chroma-based 

```{r}
"features/wednesday-w-1.json" |>                           # Change the track
  compmus_chroma(norm = "identity") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic() 
```

### Timbre-based

```{r}
"features/wednesday-w-1.json" |>                           # Change the track
  compmus_mfccs(norm = "identity") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic() 
```

Column{.tabset}
----

### Chromagram of Track 2

```{r}
"features/wednesday-w-2.json" |>                           # Change the track
  compmus_chroma(norm = "identity") |>                 
  # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(option = "G") +              
  # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL, main = "track 1 chromagrams") +
  theme_classic()  
```

### Cepstrogram

```{r}
"features/wednesday-w-2.json" |>                           #Change the track
  compmus_mfccs(norm = "identity") |>                  #Change the norm
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +               #Change the colours?
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic()                                      
```

### Chroma-based 

```{r}
"features/wednesday-w-2.json" |>                           # Change the track
  compmus_chroma(norm = "identity") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic() 
```

### Timbre-based

```{r}
"features/wednesday-w-2.json" |>                           # Change the track
  compmus_mfccs(norm = "identity") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic() 
```
