---
title: "WednesdayNo1"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    storyboard: true
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
library(shiny)
library(jsonlite)#essentia 
library(readr)
library(ggplot2)
library(tidyverse)
source("compmus.R")
compmus2025 <- read_csv("compmus2025.csv")
```
## 4ever Fixed 

# Week9 assignment page

Column{.tabset}
--------

### Chromagram of Track 1 

```{r}
"features/wednesday-w-1.json" |>                           # Change the track
  compmus_chroma(norm = "identity") |>                 
  # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +               
  # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()  
```

### Cepstrogram 

```{r}
"features/wednesday-w-1.json" |>                           #Change the track
  compmus_mfccs(norm = "identity") |>                  #Change the norm
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +               #Change the colours?
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic()                                      
```

### Chroma-based 

```{r}
"features/wednesday-w-1.json" |>                           # Change the track
  compmus_chroma(norm = "identity") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic() 
```

### Timbre-based

```{r}
"features/wednesday-w-1.json" |>                           # Change the track
  compmus_mfccs(norm = "identity") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic() 
```

Column{.tabset}
----

### Chromagram of Track 2

```{r}
"features/wednesday-w-2.json" |>                           # Change the track
  compmus_chroma(norm = "identity") |>                 
  # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +              
  # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL, main = "track 1 chromagrams") +
  theme_classic()  
```

### Cepstrogram

```{r}
"features/wednesday-w-2.json" |>                           #Change the track
  compmus_mfccs(norm = "identity") |>                  #Change the norm
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +               #Change the colours?
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic()                                      
```

### Chroma-based 

```{r}
"features/wednesday-w-2.json" |>                           # Change the track
  compmus_chroma(norm = "identity") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic() 
```

### Timbre-based

```{r}
"features/wednesday-w-2.json" |>                           # Change the track
  compmus_mfccs(norm = "identity") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic() 
```

# Introduction {.tabset}

### Origin of My Selected Tracks {.column .center data-width="10%"}

```{r}
valueBox("AIVA’s Styles Library")
```

### I was particularly drawn to their groove, rhythmic patterns and danceable energy. The uplifting beats and engaging melodies create an inspiring and refreshing listening experience. \n Curious whether there is a pattern or template that can be extracted. Specifically, a structured approach that could transform a calm or melancholic piece of music into an exciting, dance-inducing track. {.column .center data-width="50%"}

::: {style="text-align: center;"}
<p style="font-size: 18px; max-width: 80%; margin: auto;">

</p>
:::

```{r}
valueBox("Why These?")
```

### blahblahblah

```{r}
valueBox("Analysis")
```

# Hidden secret of Class Corpus

## Column {.tabset}

### How Mainstream Are we?

```{r}
boxplot(
  compmus2025$approachability,
  horizontal = TRUE,
  col = "lightblue",
  main = "Approachability Boxplot",
  xlab = "Approachability Score",
  outpch = 16,
  outcol = "lightpink"
)
```

## Column {.tabset}

### How Mainstream Are we?

```{r}
hist(compmus2025$approachability, 
     breaks = 10, 
     col = "skyblue",
     main = "Distribution of Approachability",
     xlab = "Approachability Score",
     ylab = "Frequency")
```

## Column {.tabset}

### How Mainstream Are we?

```{r}
boxplot(
  compmus2025$approachability,
  horizontal = TRUE,
  col = "lightblue",
  main = "Approachability Boxplot",
  xlab = "Approachability Score",
  outpch = 16,
  outcol = "lightpink"
)
```

# Visualisation

## Column {.tabset}

### Chromagram

```{r}
"features/wednesday-w-1.json" |>                           # Change the track
  compmus_chroma(norm = "identity") |>                 
  # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +              
  # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()  
```

### Cepstrogram

```{r}
"features/wednesday-w-1.json" |>                           #Change the track
  compmus_mfccs(norm = "identity") |>                  #Change the norm
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +               #Change the colours?
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic()                                      
```

### Tempo vs Danceablity

```{r fig.height=4, fig.width=6}
compmus2025 |> 
  ggplot(aes(x = tempo,y=danceability)) + 
  geom_jitter() + geom_smooth() -> p1
ggplotly(p1)
```

> **Analysis:**<br><br> - Danceability is highest at 60-80 BPM, suggesting that slower-paced music might be more suitable for dancing. <br><br> - Danceability significantly decreases above 130 BPM, indicating that fast-paced music is more suitable for listening rather than dancing.

## Column {.tabset}

### Tempo vs Arousal

```{r}
"features/sytze-m-1.json" |>                           #Change the track
  compmus_mfccs(norm = "identity") |>                  #Change the norm
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +               #Change the colours?
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic()                                      
```

> **Analysis:**<br><br> - Music with tempos between 100-130 BPM has the highest Arousal, suggesting that this tempo range makes music more exciting and stimulating, making it ideal for boosting attention or enhancing emotions.<br><br> - Arousal starts to decrease above 130 BPM, possibly because the extremely fast tempo causes music to lose emotional expression, making it feel more monotonous or mechanical.

### +Danceablity

```{r, fig.width=6, fig.height=4}
compmus2025 |>                   # Start with the data
  ggplot(                     # Set up the plot.
    aes(
      x = tempo,
      y = arousal,
      colour = danceability
    )
  ) +geom_point() +              # Scatter plot.
scale_x_continuous(         # 设置x范围
    limits = c(60, 160),
    breaks = c(60,80,100,120,140, 160), # Specify grid lines
    minor_breaks = NULL       # Remove 'minor' grid lines.
  ) +
  scale_y_continuous(         # Fine-tune the y axis in the same way.
    limits = c(4, 5),
    breaks = c(3, 6),
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c() +  # 颜色渐变
  scale_size_continuous(      # Fine-tune the sizes of each point.
    trans = "exp",            # Use an exp transformation to emphasise loud..
    guide = guide_colorbar(barwidth = 10)        # Remove the legend for size.
  ) +
  theme_light() +             # Use a simpler theme.
  labs(                       # Make the titles nice.
    x = "Tempo",
    y = "Arousal",
    colour = "Danceability  "
  )+
  theme(
    legend.key.width = unit(1, "cm"),  # **固定图例宽度**
    legend.key.height = unit(1, "cm")  # **固定图例高度**
  )
```

### +Valence

```{r, fig.width=6, fig.height=4}
ggplot( compmus2025,
      aes(x = tempo,
      y = arousal,
      colour = valence
    )
  ) +geom_point() +              # Scatter plot.
scale_x_continuous(         # 设置x范围
    limits = c(60, 160),
    breaks = c(60,80,100,120,140, 160), # Specify grid lines
    minor_breaks = NULL       # Remove 'minor' grid lines.
  ) +
  scale_y_continuous(         # Fine-tune the y axis in the same way.
   limits = c(4, 5),
    breaks = c(3, 6),
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c() +  # 颜色渐变
  scale_size_continuous(      # Fine-tune the sizes of each point.
    trans = "exp",            # Use an exp transformation to emphasise loud..
    guide = guide_colorbar(barwidth = 10)          # Remove the legend for size.
  ) +
  theme_light() +             # Use a simpler theme.
  labs(                       # Make the titles nice.
    x = "Tempo",
    y = "Arousal",
    colour = "Valence Score"
  )+
  theme(
    legend.key.width = unit(1, "cm"),  # **固定图例宽度**
    legend.key.height = unit(1, "cm")  # **固定图例高度**
  )
```

```{r, eval=FALSE}
fig.width=10, fig.height=5
> captions with inline code
```

# Conclusion
