---
title: "WednesdayNo1"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    storyboard: true
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
library(shiny)
library(jsonlite)#essentia 
```

# Introduction{.tabset}

### Origin of My Selected Tracks{.column data-width=10% .center}

```{r}
valueBox("AIVA’s Styles Library")
```
### I was particularly drawn to their groove, rhythmic patterns and danceable energy. The uplifting beats and engaging melodies create an inspiring and refreshing listening experience. \n Curious whether there is a pattern or template that can be extracted. Specifically, a structured approach that could transform a calm or melancholic piece of music into an exciting, dance-inducing track.{.column data-width=50% .center}
<div style="text-align: center;">
  <p style="font-size: 18px; max-width: 80%; margin: auto;">
  </p>
</div>
```{r}
valueBox("Why These?")
```
### blahblahblah
```{r}
valueBox("Analysis")
```

# Visualisation

### Plot A

```{r, eval=FALSE}
bpm_data <- data.frame(
    track = c("Track A", "Track B", "Track C"),
    bpm = c(essentia_features$rhythm.bpm)
)
```

```{r, eval=FALSE}
fig.width=10, fig.height=5
> captions with inline code
```

***

commentary zheli 

# Conclusion