---
title: "WednesdayNo1"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    storyboard: true
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
library(shiny)
library(jsonlite)#essentia 
library(readr)
library(ggplot2)
aisc2024 <- read_csv("aisc2024.csv")
```

# Introduction {.tabset}

### Origin of My Selected Tracks {.column .center data-width="10%"}

```{r}
valueBox("AIVA’s Styles Library")
```

### I was particularly drawn to their groove, rhythmic patterns and danceable energy. The uplifting beats and engaging melodies create an inspiring and refreshing listening experience. \n Curious whether there is a pattern or template that can be extracted. Specifically, a structured approach that could transform a calm or melancholic piece of music into an exciting, dance-inducing track. {.column .center data-width="50%"}

::: {style="text-align: center;"}
<p style="font-size: 18px; max-width: 80%; margin: auto;">

</p>
:::

```{r}
valueBox("Why These?")
```

### blahblahblah

```{r}
valueBox("Analysis")
```

# Visualisation

## Column {.tabset}

### Tempo vs Danceablity

```{r fig.height=4, fig.width=6}
aisc2024 |> 
  ggplot(aes(x = tempo,y=danceability)) + 
  geom_jitter() + geom_smooth() -> p1
ggplotly(p1)
```
> **Analysis:**<br><br>
- Danceability is highest at 60-80 BPM, suggesting that slower-paced music might be more suitable for dancing. <br><br>
- Danceability significantly decreases above 130 BPM, indicating that fast-paced music is more suitable for listening rather than dancing.

## Column {.tabset}

### Tempo vs Arousal

```{r, fig.width=6, fig.height=4}
#在散点图上添加随机抖动（Jitter）的函数。
#它的作用是减少点重叠（Overlapping），提高可读性。
aisc2024 |> ggplot(aes(x = tempo, y = arousal)) + 
  geom_jitter() + geom_smooth() -> p1
ggplotly(p1)
```

> **Analysis:**<br><br>
- Music with tempos between 100-130 BPM has the highest Arousal, suggesting that this tempo range makes music more exciting and stimulating, making it ideal for boosting attention or enhancing emotions.<br><br>
- Arousal starts to decrease above 130 BPM, possibly because the extremely fast tempo causes music to lose emotional expression, making it feel more monotonous or mechanical.

### +Danceablity

```{r, fig.width=6, fig.height=4}
aisc2024 |>                   # Start with the data
  ggplot(                     # Set up the plot.
    aes(
      x = tempo,
      y = arousal,
      colour = danceability
    )
  ) +geom_point() +              # Scatter plot.
scale_x_continuous(         # 设置x范围
    limits = c(60, 160),
    breaks = c(60,80,100,120,140, 160), # Specify grid lines
    minor_breaks = NULL       # Remove 'minor' grid lines.
  ) +
  scale_y_continuous(         # Fine-tune the y axis in the same way.
    limits = c(4, 5),
    breaks = c(3, 6),
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c() +  # 颜色渐变
  scale_size_continuous(      # Fine-tune the sizes of each point.
    trans = "exp",            # Use an exp transformation to emphasise loud..
    guide = guide_colorbar(barwidth = 10)        # Remove the legend for size.
  ) +
  theme_light() +             # Use a simpler theme.
  labs(                       # Make the titles nice.
    x = "Tempo",
    y = "Arousal",
    colour = "Danceability  "
  )+
  theme(
    legend.key.width = unit(1, "cm"),  # **固定图例宽度**
    legend.key.height = unit(1, "cm")  # **固定图例高度**
  )
```

### +Valence

```{r, fig.width=6, fig.height=4}
ggplot( aisc2024,
      aes(x = tempo,
      y = arousal,
      colour = valence
    )
  ) +geom_point() +              # Scatter plot.
scale_x_continuous(         # 设置x范围
    limits = c(60, 160),
    breaks = c(60,80,100,120,140, 160), # Specify grid lines
    minor_breaks = NULL       # Remove 'minor' grid lines.
  ) +
  scale_y_continuous(         # Fine-tune the y axis in the same way.
   limits = c(4, 5),
    breaks = c(3, 6),
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c() +  # 颜色渐变
  scale_size_continuous(      # Fine-tune the sizes of each point.
    trans = "exp",            # Use an exp transformation to emphasise loud..
    guide = guide_colorbar(barwidth = 10)          # Remove the legend for size.
  ) +
  theme_light() +             # Use a simpler theme.
  labs(                       # Make the titles nice.
    x = "Tempo",
    y = "Arousal",
    colour = "Valence Score"
  )+
  theme(
    legend.key.width = unit(1, "cm"),  # **固定图例宽度**
    legend.key.height = unit(1, "cm")  # **固定图例高度**
  )
```

```{r, eval=FALSE}
fig.width=10, fig.height=5
> captions with inline code
```

# Conclusion
