---
title: "Make the World Dance"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    storyboard: true
    theme: 
      version: 4
      bootswatch: minty
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(tidymodels)
library(ggdendro)
library(heatmaply)
library(plotly)
library(shiny)
library(jsonlite) 
library(readr)
library(ggplot2)
library(tidyverse)
source("compmus.R")
compmus2025 <- read_csv("compmus2025.csv")
#      C     C#    D     Eb    E     F     F#    G     Ab    A     Bb    B
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )
get_conf_mat <- function(fit) {
  outcome <- .get_tune_outcome_names(fit)
  fit |> 
    collect_predictions() |> 
    conf_mat(truth = outcome, estimate = .pred_class)
}  

get_pr <- function(fit) {
  fit |> 
    conf_mat_resampled() |> 
    group_by(Prediction) |> mutate(precision = Freq / sum(Freq)) |> 
    group_by(Truth) |> mutate(recall = Freq / sum(Freq)) |> 
    ungroup() |> filter(Prediction == Truth) |> 
    select(class = Prediction, precision, recall)
}  
cluster_juice <-
  recipe(
    filename ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = compmus2025
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |> 
  # step_range(all_predictors()) |> 
  prep(compmus2025) |>
  juice() |>
  column_to_rownames("filename")
compmus_dist <- dist(cluster_juice, method = "euclidean")
```

# Introduction {.tabset}

## ✨ What Makes Music Danceable?
_Journeying through groove, emotion, and structure_

---

### 🔹 Finding 1: **Speed Isn’t Everything**

Even though faster songs seem more energetic, **tempo alone doesn't guarantee danceability**.  
In my analysis, `wednesday-w-2` had a higher BPM than `roemer-i-2`, but was outperformed in danceability.  
> *A fast beat might excite the body, but not necessarily move the feet.*

---

### 🔹 Finding 2: **Emotion Drives Movement**

The top 5 most danceable tracks had:

- **Higher arousal** (emotional energy)  
- **Higher valence** (positive mood)  
- **Lower instrumentalness** (more vocal presence)

This suggests that emotional intensity and human-like qualities are key to making music feel danceable.

---

### 🔹 Finding 3: **Structure Matters More Than You Think**

From chroma-based self-similarity and cepstrograms, I found:

- Danceable tracks often show **clear repetition** and **rhythmic regularity**.
- These mid-level structures (groove, loops, call-and-response) are not captured by summary features alone.

> *Danceability lives in the structure, not just in the stats.*

---

### 🔹 Finding 4: **Familiarity ≠ Danceability**

Despite assumptions, **approachability** (how familiar or “safe” a track feels) showed **no strong correlation** with danceability.  
> *Being predictable doesn’t mean it makes you move.*

---

## 🎯 Final Takeaway

> **Danceability is an emergent property.**  
> It doesn’t come from a single number.  
> It emerges from the interaction between **rhythm, energy, emotion, and structure over time**.

By zooming in on my own tracks with chromagrams, MFCCs, self-similarity, and chordograms, I realized:  
🎶 *Danceable music is not just heard — it’s felt in your bones.*


### Origin of My Selected Tracks. I sourced my tracks from this AI Styles Library—a surprisingly expressive and diverse set of tracks generated by an AI trained in music composition. This made it the perfect starting point to explore what makes music danceable. {.column .center data-width="10%"}

```{r}
valueBox("AIVA’s Styles Library")
```

### I was particularly drawn to their groove, rhythmic patterns and danceable energy. The uplifting beats and engaging melodies create an inspiring and refreshing listening experience. \n Curious whether there is a pattern or template that can be extracted. Specifically, a structured approach that could transform a calm or melancholic piece of music into an exciting, dance-inducing track. {.column .center data-width="50%"}

::: {style="text-align: center;"}
<p style="font-size: 18px; max-width: 80%; margin: auto;">

</p>
:::

```{r}
valueBox("Why These?")
```

### 首先分别提取比较danceability高和低的前三之间各种tracklevel的区别？3，预测时可以分别看ai和nonai根据预测方法的准确性

```{r}
valueBox("Analysis")
```

### Top 5 Danceability {.column}

```{r}
compmus_dist |> 
  hclust(method = "single") |> # Try single, average, and complete.
  dendro_data() |>
  ggdendrogram()
top5 <- cluster_juice |> arrange(desc(danceability)) |> head(5)
bottom5 <- cluster_juice |> arrange(danceability) |> head(5)
heatmaply(
  top5,
  main = "Top 5 Danceability Tracks",
  dist_method = "euclidean",
  hclust_method = "average"
)
```

### Bottom 5 Danceability {.column}

```{r}
heatmaply(
  bottom5,
  main = "Bottom 5 Danceability Tracks",
  dist_method = "euclidean",
  hclust_method = "average"
)
```

# 1st Attempt at Explaining Danceability--Corpus-level comparison

## Column {.tabset}

### Hypothesis 1: Tempo → Danceability

```{r}
aisc2024 <- compmus2025 |> 
  mutate(
    highest =	filename %in% c("roemer-i-1", "roemer-i-2"),
    is_mine = filename %in% c("wednesday-w-1", "wednesday-w-2"),
    label = case_when(
      is_mine ~ filename,
      highest ~ "roemer-i-2",   # 只给 roemer-i-2 加标签
      TRUE ~ NA_character_
    )
  )

# 画图
ggplot(
  aisc2024,
  aes(
    x = tempo,
    y = danceability,
    colour = is_mine | highest
  )
) +
  geom_point(alpha = 0.6) +
  geom_text(aes(label = label), hjust = -0.1, vjust = 0.3, size = 3, na.rm = TRUE) +
  scale_color_manual(
    values = c("FALSE" = "grey80", "TRUE" = "darkgreen"),
    labels = c("Other Tracks", "Selected Tracks"),
    name = "Track Type"
  ) +
  scale_x_continuous(
    limits = c(50, 200),
    breaks = c(50, 100, 150, 180),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    minor_breaks = NULL
  ) +
  scale_size_continuous(trans = "exp", guide = "none") +
  theme_light() +
  labs(
    title = "Where My Tracks Sit in the Danceability Space",
    x = "Tempo (BPM)",
    y = "Danceability",
    colour = "Track Type"
  )
```