---
title: "Make the World Dance"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    storyboard: true
    theme: 
      version: 4
      bootswatch: minty
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(tidymodels)
library(ggdendro)
library(heatmaply)
library(plotly)
library(shiny)
library(jsonlite) 
library(readr)
library(ggplot2)
library(tidyverse)
source("compmus.R")
compmus2025 <- read_csv("compmus2025.csv")
#      C     C#    D     Eb    E     F     F#    G     Ab    A     Bb    B
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )
get_conf_mat <- function(fit) {
  outcome <- .get_tune_outcome_names(fit)
  fit |> 
    collect_predictions() |> 
    conf_mat(truth = outcome, estimate = .pred_class)
}  

get_pr <- function(fit) {
  fit |> 
    conf_mat_resampled() |> 
    group_by(Prediction) |> mutate(precision = Freq / sum(Freq)) |> 
    group_by(Truth) |> mutate(recall = Freq / sum(Freq)) |> 
    ungroup() |> filter(Prediction == Truth) |> 
    select(class = Prediction, precision, recall)
}  
cluster_juice <-
  recipe(
    filename ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = compmus2025
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |> 
  # step_range(all_predictors()) |> 
  prep(compmus2025) |>
  juice() |>
  column_to_rownames("filename")
compmus_dist <- dist(cluster_juice, method = "euclidean")
```

# Introduction {.tabset}

### Origin of My Selected Tracks. I sourced my tracks from this AI Styles Library—a surprisingly expressive and diverse set of tracks generated by an AI trained in music composition. This made it the perfect starting point to explore what makes music danceable. {.column .center data-width="10%"}

```{r}
valueBox("AIVA’s Styles Library")
```

### I was particularly drawn to their groove, rhythmic patterns and danceable energy. The uplifting beats and engaging melodies create an inspiring and refreshing listening experience. \n Curious whether there is a pattern or template that can be extracted. Specifically, a structured approach that could transform a calm or melancholic piece of music into an exciting, dance-inducing track. {.column .center data-width="50%"}

::: {style="text-align: center;"}
<p style="font-size: 18px; max-width: 80%; margin: auto;">

</p>
:::

```{r}
valueBox("Why These?")
```

### 首先分别提取比较danceability高和低的前三之间各种tracklevel的区别？3，预测时可以分别看ai和nonai根据预测方法的准确性

```{r}
valueBox("Analysis")
```

### Top 5 Danceability {.column}

```{r}
compmus_dist |> 
  hclust(method = "single") |> # Try single, average, and complete.
  dendro_data() |>
  ggdendrogram()
top5 <- cluster_juice |> arrange(desc(danceability)) |> head(5)
bottom5 <- cluster_juice |> arrange(danceability) |> head(5)
heatmaply(
  top5,
  main = "Top 5 Danceability Tracks",
  dist_method = "euclidean",
  hclust_method = "average"
)
```

### Bottom 5 Danceability {.column}

```{r}
heatmaply(
  bottom5,
  main = "Bottom 5 Danceability Tracks",
  dist_method = "euclidean",
  hclust_method = "average"
)
```

# 1st Attempt at Explaining Danceability--Corpus-level comparison

## Column {.tabset}

### Hypothesis 1: Tempo → Danceability

```{r}
aisc2024 <- compmus2025 |> 
  mutate(
    highest =	filename %in% c("roemer-i-1", "roemer-i-2"),
    is_mine = filename %in% c("wednesday-w-1", "wednesday-w-2"),
    label = case_when(
      is_mine ~ filename,
      highest ~ "roemer-i-2",   # 只给 roemer-i-2 加标签
      TRUE ~ NA_character_
    )
  )

# 画图
ggplot(
  aisc2024,
  aes(
    x = tempo,
    y = danceability,
    colour = is_mine | highest
  )
) +
  geom_point(alpha = 0.6) +
  geom_text(aes(label = label), hjust = -0.1, vjust = 0.3, size = 3, na.rm = TRUE) +
  scale_color_manual(
    values = c("FALSE" = "grey80", "TRUE" = "lightgreen"),
    labels = c("Other Tracks", "Selected Tracks"),
    name = "Track Type"
  ) +
  scale_x_continuous(
    limits = c(50, 200),
    breaks = c(50, 100, 150, 180),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    minor_breaks = NULL
  ) +
  scale_size_continuous(trans = "exp", guide = "none") +
  theme_light() +
  labs(
    title = "Where My Tracks Sit in the Danceability Space",
    x = "Tempo (BPM)",
    y = "Danceability",
    colour = "Track Type"
  )
```